### 2020-09-16 tables/kinematictable.tex

\begin{deluxetable}{cccccccccccc}
\rotate
\tablewidth{0pt}
%\rotate
\tabletypesize{\scriptsize}
\tablecaption{Kinematic \pdspy Modeling}
\tablehead{
  \colhead{Source} & \colhead{RA Offset} & \colhead{Dec Offset} & \colhead{Inc.} & \colhead{P.A.} & \colhead{M$_{*}$} & \colhead{M$_{gas}$\tablenotemark{a}} &\colhead{R$_{disk}$} & \colhead{V$_{sys}$}& \colhead{Turbulence}& \colhead{Surface Density Index $\gamma$} & \colhead{T$_{0}$}\\
  & \colhead{(\arcsec)} & \colhead{(\arcsec)} & \colhead{(\deg)} & \colhead{(\deg)} & \colhead{\solm} & \colhead{\solm}& \colhead{(au)} & \colhead{(km~s$^{-1}$)}& \colhead{(km~s$^{-1}$)}& & \colhead{(K)}  \\
}
\startdata
IRS3B  & $0.031^{+0.019}_{-0.011}$ & $0.025^{+0.020}_{-0.015}$ & $66.0^{+ 3.0}_{- 4.6} $ & $26.7^{+  1.8}_{-  2.9} $ & $1.19^{+0.13}_{-0.07}$ & $0.079^{+0.021}_{-0.016}$ & $299.0^{+ 24.9}_{- 47.6}$ & $4.880^{+0.110}_{-0.090}$ & $0.012^{+ 0.005}_{- 0.003}$ & $  1.2^{+  0.1}_{-  0.1}$&$   50^{+    3}_{-    5}$\\
IRS3A  & $0.034^{+0.003}_{-0.003}$ & $0.015^{+0.003}_{-0.003}$ & $69.5^{0.38}_{0.37} $  & $122.4^{1.4}_{1.4}$ & $1.51^{+0.06}_{-0.07}$ & $6.3^{+1.6}_{-1.3}x10^{-6}$ & $ 39.9^{+  2.4}_{-  1.4}$ & $5.288^{+0.090}_{-0.084}$ & $0.015^{+ 0.06}_{-0.009}$ & $0.4^{+0.2}_{-0.1}$ & $163^{+9}_{-8}$ \\
% /home/reynolds/local/PDSPY_models/L1448IRS3A/h13cn/localruns/
% /home/reynolds/lovell1/c17o
\enddata
\tablecomments{Summary of kinematic model parameters. The RA and DEC offsets of the \pdspy modeling are defined from the central positions given in PV Analysis, Table~\ref{table:pvtable}. The errors presented are the 3-$\sigma$\space confidence intervals of the best fit walkers generated from \textit{emcee}.}
\tablenotetext{a}{\added{The reported values of M$_{gas}$\space depend on the assumed abundance for each of the molecules. For the IRS3B source, we used the \cso\space emission, which has an assumed abundance of $2\times10^{-7}$, while for the IRS3A source we used the \htcn\space emission which has an assumed abundance of $2.9\times10^{-11}$.}}
\end{deluxetable}\label{table:pdspykinematic}

% /net/ryle/myhome1/reynolds/PDSPY_models/oscer/wide/h13cn
% ~/jansky1/c17o
% irs3bcont
% irs3acont